---
title: "PNInputEM"
author: "Gregory Jefferis"
date: "27/06/2017"
output: html_document
---

# Intro

We're going to take a look at the inputs that Alex has found for the neurons
marked up as "Mike's C11"

```{r startup, message=FALSE}
library(elmr)
library(dplyr)
rgl::setupKnitr()
```


```{r, cache=TRUE}
mikec11=read.neurons.catmaid("annotation:^PD2a1/b1")
```

```{r, cache=TRUE}
pncc = catmaid_get_connectors_between('annotation:^PN$', 
                                    post_skids = "annotation:^PD2a1/b1")
```

Where are the inputs located? For each synapse, see if it's close to MB or LH
surface (a few outputs are closer to LH).

```{r}
lhr=as.mesh3d(subset(FAFB13NP.surf, "LH_R"))
mbr=as.mesh3d(subset(FAFB13NP.surf, "MB_CA_R"))

pncc %>% 
  select(connector_x:connector_z) %>%
  pointsinside(surf = lhr, rval = 'dist') -> lhdist
pncc %>% 
  select(connector_x:connector_z) %>%
  pointsinside(surf = mbr, rval = 'dist') -> mbdist
pncc$np=ifelse(mbdist>lhdist, "MB", "LH")
```

OK, let's summarise synapse location. Quite a few 

```{r}
table(pncc$np)
table(pncc$post_skid, pncc$np)
```

Let's define some functions to handle PN names.

```{r}
get_names <- function(x) {
  if(is.numeric(x)) catmaid_get_neuronnames(x) else x
}
glom <- function(x) {
  nn=get_names(x)
  stringr::str_match(nn, "[Gg]lomerulus ([A-z1-7]{1,}[^ ]*) ")[,2]
}
```

```{r}
table(pncc$np, glom(pncc$pre_skid), pncc$post_skid)
```

Now let's take a look at all connections onto the dendrites of LHNs

First fetch all connections

```{r}
incc = catmaid_get_connectors_between(post_skids = "annotation:^PD2a1/b1")
```

Now assign them to LH, Calyx or Other
```{r}
# less than 5 µm outside LH surface
inlh=pointsinside(incc[,c('connector_x', 'connector_y','connector_z')],
                       surf = lhr, rval = 'dist') > -5e3

inmb=pointsinside(incc[,c('connector_x', 'connector_y','connector_z')],
                       surf = mbr, rval = 'dist') > -5e3
incc$np='other'
incc$np[inlh]='LH'
incc$np[inmb]='MB'
table(incc$np)
```

```{r, webgl=TRUE}
clear3d()
points3d(incc[,c('post_node_x', 'post_node_y','post_node_z')],
         col=rainbow(3)[factor(incc$np)])
wire3d(lhr)
```

OK now that we have divided inputs, let's look at fraction of them that are 
PNs etc

```{r}
pn.ids=catmaid_skids('annotation:^PN$')
lhdone.ids=catmaid_skids('annotation:^LH_DONE$')
pn.done.ids=intersect(pn.ids, lhdone.ids)
a2sc.ids=catmaid_skids("annotation:^MBON a2sc")
```

```{r}
incc$intype='other'
incc$intype[incc$pre_skid%in%pn.ids]='PN'
incc$intype[incc$pre_skid%in%a2sc.ids]='a2sc'
```

Let's cross-classify inputs by presynaptic type and location:

```{r}
xt=xtabs(~intype+np, data = incc)
xt
```

So `r xt['a2sc','LH']/sum(xt[c('a2sc','PN'),'LH'])*100` of LH inputs are from
the a2sc.

Let's look at the relationship between input type and location

```{r, webgl=TRUE}
clear3d()
points3d(incc[,c('post_node_x', 'post_node_y','post_node_z')],
         col=rainbow(3)[factor(incc$intype)])
wire3d(lhr)
```

Might want to check a few of those a2sc inputs for dendritic vs axonal.

Summarise by LHON

```{r}
incc %>%
  filter(np=='LH') %>%
  xtabs(~post_skid+intype, data=.)
```


```{r}
incc %>%
  xtabs(~post_skid+intype+np, data=.)
```


## Unidentified inputs

For the unidentified inputs, would be nice to get an indication of 
whether they are attached to big neurons or not. First for the MB.

```{r MB unidentified}
subset(incc, np=='MB' & intype!='PN')[,'pre_skid'] -> mbinputs
nn=catmaid_get_neuronnames(mbinputs) 
data.frame(skid=names(nn), name=nn)  %>% 
  group_by(skid, name) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) -> sdf
knitr::kable(sdf)
```

Now for the LH

```{r LH unidentified}
subset(incc, np=='LH' & intype!='PN')[,'pre_skid'] -> lhinputs
lhnn=catmaid_get_neuronnames(lhinputs) 
data.frame(skid=names(lhnn), name=lhnn)  %>% 
  group_by(skid, name) %>% 
  summarise(n=n()) %>% 
  arrange(desc(n)) -> sdflh
plot(table(sdflh$n))
plot(ecdf(sdflh$n), verticals = T, 
     main = 'Cumulative Distribution Synaptic Partner Strength',
     ylab='Cumulative Fraction Total Inputs',
     xlab='Connections per partner')
# just take a look at inputs n>1
knitr::kable(filter(sdflh, n>1))
```


## Input clustering

Let's see if there is different classes of inputs are located on different
parts of the LHN dendritic tree. We could ask a few different questions.

* Is there any relationship between Strahler order and input type?
* Are inputs of the same type closer together by geodesic or Euclidean distance?
  * I think geodesic might be more interesting and less obvious

Let's start by taking a looking using LHN with most inputs

```{r, rgl=TRUE}
with(subset(incc, post_skid==1299700 & pre_skid %in% a2sc.ids),
     spheres3d(cbind(connector_x, connector_y, connector_z), 
               radius=300, col='green'))
```

### Strahler order
First calculate the strahler order for the connections
```{r}
incc$so=NA_integer_
for(n in mikec11){
  so=strahler_order(n)
  
  # which of the nodes in this neuron have synaptic input
  post_node_idxs=which(n$d$PointNo%in%incc$post_node_id)
  incc$so[match(n$d$PointNo[post_node_idxs],incc$post_node_id)]=
    so$points[post_node_idxs]
}
```

Looking at these 

```{r}
xx=xtabs(~so+I(intype=='PN'), data=subset(incc, np=="LH" & intype%in%c("PN","other")))
xx
summary(xx)
xx=xtabs(~so+I(intype=='PN'), data=subset(incc, np=="LH" & intype%in%c("PN","a2sc")))
xx
summary(xx)
```

The conclusion is that there are some small asymmetries in the distribution but 
the effect sizes appear fairly small.

```{r, eval=FALSE}
incc %>%
  filter(np=='LH' & intype=='other') %>%
  xtabs(~post_skid+)
```

